Lambda Calculus through Tropical Mathematics

### 2023-01-12 section6.tex
commit

FOLLOW SECTION III.C for definitions of models of LL and DiLL.

@@

Also in this case, the construction above generalizes the situation in $\LREL$: for the modules of the form $\Lawv^{X}$, the symmetric algebra construction coincides with the usual one, provided by multisets.
\begin{proposition}
For any set $X$, 
$!_{n}(\Lawv^{X})\simeq \Lawv^{\C M_{n}(X)}$, where
$\C M_{n}(X)$ is the set of multisets of $X$ of cardinality $n$. 
\end{proposition}

At this point, we can define the exponential comonad as 
$!M:= \prod_{n\in \BB N}!_{n}M$, obtaining a Lafont category.
ECC....

>>>>

First, we need to define a Lafont exponential $!$ over $\Mod$, and for this we use a well-known recipe from \cite{Mellies, Weight}, that is: we first define a symmetric algebra $\Sym_{n}(M)$ as the equalizer of all permutative actions on $n$-tensors $M\otimes \dots \otimes M$; then, exploiting the fact that tensors and products commute in $\Mod$, we define $!$ as the product of the symmetric algebras $!_{n}$.

@@

Using the fact that $\prod_{i}M_{i}\otimes  N\simeq (\prod_{i}M_{i})\otimes N$ holds in $\Mod$ [\cite{}, p.~???], we obtain the following:
\begin{theorem}
The functor $!M:= \prod_{n\in \BB N}!_{n}M$ is a Lafont exponential in $\Mod$.
Hence, $\Mod_{!}$ (equivalently, $\GMet_{!}$) is cartesian closed. 
\end{theorem}

Also in this case of exponentials, the constructions above generalize the situation in $\LREL$:
\begin{proposition}
For any set $X$, 
$!_{n}(\Lawv^{X})\simeq \Lawv^{\C M_{n}(X)}$, where
$\C M_{n}(X)$ is the set of multisets of $X$ of cardinality $n$, and 
$!(\Lawv^{X})\simeq \Lawv^{\multiset(X)}$. 
\end{proposition}

To conclude, we mush show that $\Mod$ is actually a model of $\STDLC$, that is, a cartesian closed differential category:
\begin{enumerate}
\item $\Mod$ is a left-additive-CCC, since each hom-set is a commutative monoid with respect to $\infty$ and $\min$, and the cartesian closed structure is well-behaved with respect to that;

\item $\Mod$ can be equipped with a differential operator defined, for $f:!M\to N$, by 
\begin{align}\label{eq:dermod}
\Der[f](\alpha)=
\bigvee\left\{
f(\beta\cup [x]) \ \Big \vert  \ 
\iota_{n}(\beta)\otimes \iota_{1}(x) \leq S(\alpha)
\right\}
\end{align}
where $\iota_{k}: M_{k}\to \prod_{i\in I}M_{i}$ is the injection morphism given by $\iota_{k}(x)( k)=x$ and $\iota_{k}(x)(i\neq k)=\infty$,
and $S: !(X\times Y)\to !X\otimes !Y$ is the Seely isomorphism \cite{}, 
satisfying all axioms (D1)-(D7) and (DCurry), see the Appendix.
\end{enumerate}

This leads then to:
\begin{theorem}
$\Mod_{!}$ (equivalently, $\GMet_{!}$), equipped with $\Der$, is a $CC\partial C$.
\end{theorem}

Notice that, when $f:!(\Lawv^{X})\to \Lawv^{Y}$, so that $f$ is described by a matrix $\widehat f: \multiset(X)\times Y\to \Lawv$, one can check that \eqref{eq:dermod} coincides with the definition of the differential operator in $\LREL$.
